---
output:
  word_document: default
  html_document: default
---
                    BitCoin Regression Analysis using Principla Components 

```{r}
library(corrplot)
library(ggplot2)
library(ggfortify)
library(zoo)
library(dplyr)
library(lubridate)
library(plotly)

```

```{r}
bitcoin <- read.csv("bitcoin_data.csv", header = TRUE)
str(bitcoin)
summary(bitcoin)

bitcoin$Date <- as.Date(bitcoin$Date, format = "%m/%d/%Y")
str(bitcoin)
index <- which(is.na(bitcoin$trade_volume))
summary(index)
bitcoin[is.na(bitcoin)] = 0
```
Trend plays an important role in trading, since there was no or few hundreds of hike in Bitcoin price from 2009 till 2015 
```{r}
hist(bitcoin$Price, col = "skyblue")
mybitcoin <- subset(bitcoin, bitcoin$Date >= '0016-06-01')
head(mybitcoin)

```


Bitcoin price  trend over the period of time 
```{r}
ggplot(bitcoin, aes(bitcoin$Date, bitcoin$Price,)) + geom_line() +
  geom_point(color = "dark blue") +
  ggtitle("Bitcoin values in USD trend") +
  labs ( x="Date" , y="Bitcoin price in USD") 
```
Correlation on Market Price 
```{r}
cor = cor(mybitcoin[,c(2:13)])
corrplot(cor, method = "square", type="upper", tl.srt = 65, tl.col = "black", tl.cex = 0.7, 
title = "")

#list of highly corelated fields 
correlation = as.data.frame(cor)
highly_correlated = correlation[correlation$market_price>= 0.75, ]
rownames(highly_correlated)


library(PerformanceAnalytics)
cor <- mybitcoin[, c(2:13)]
chart.Correlation(cor, histogram=TRUE, pch=19)
```
# Check if there are any missing values
```{r}
library(Rcpp)
library(Amelia)
missing<-missmap(mybitcoin[2:13], main = "Missing values vs observed")
summary(mybitcoin$trade_volume)
```
We are not using column Date for our further regression model.

```{r}
mybitcoin <- select(mybitcoin, -Date)
```

```{r}

initial_regression <- lm(Price ~ . , data=mybitcoin)
summary(initial_regression)

library(ggplot2)
library(ggfortify)
library(car)
autoplot(initial_regression, col = "goldenrod1") +  theme_minimal()
#residualPlots(initial_regression)

#Test for constant variance fails
lmtest::bptest(initial_regression) # Breusch-Pagan test
car::ncvTest(initial_regression)  # Breusch-Pagan test
shapiro.test(residuals(initial_regression))

```
All variables are significant and R ^2 is 99.47%, which seems to be overfitting 

```{r}
pcabitcoin=mybitcoin[,-3,13]

```

2. PCA

```{r}
pca_bitcoin <- prcomp(scale(pcabitcoin[,-c(12,13,14)])) # removing Date, response
summary(pca_bitcoin)

plot(pca_bitcoin)
autoplot(pca_bitcoin, col= "blue")

biplot(pca_bitcoin)

```
```{r}
# sqrt of eigenvalues
#pca_bitcoin$sdev
pca_bitcoin$rotation
```
                              PC1         PC2         PC3         PC4          PC5
Day                          0.10839617 -0.37296518 -0.57861384  0.03483389  0.070543539
Week                         0.11183986 -0.36857620 -0.58202769  0.03536181  0.065073134
Transaction_fees..BTC.      -0.18412476 -0.41741116  0.31605826 -0.09918104 -0.207288582
No_of_transactions          -0.01628881 -0.44004318  0.27801378 -0.54949391  0.503308158
Output_value..BTC.           0.13179510 -0.33782768  0.31571415  0.80893038  0.281300888
estimated_transaction_value -0.38585200 -0.25619785  0.02779089  0.06961858 -0.017386038
Miners_revenue              -0.41647402 -0.16889134 -0.01041179 -0.02157734  0.040479470
Cost_per_transaction        -0.43426583  0.01981957 -0.07564301  0.10202948 -0.002831797
Difficulty                  -0.37167268  0.24899477 -0.15200716  0.08695035  0.357914898
Hash_rate                   -0.38058192  0.23533726 -0.13512286  0.07029819  0.389944922
Trade_volume                -0.36097444 -0.17639295 -0.04028549  0.05623887 -0.577288707

```{r}
#As per PCA importance 
fit.pca <- lm(Price ~ estimated_transaction_value+Difficulty+Hash_rate+Cost_per_transaction+Miners_revenue+Trade_volume,  data=mybitcoin)
summary(fit.pca)

autoplot(fit.pca, col = "goldenrod1") +  theme_minimal()
residualPlots(fit.pca)

#Test for constant variance fails
lmtest::bptest(fit.pca) # Breusch-Pagan test
car::ncvTest(fit.pca)  # Breusch-Pagan test
shapiro.test(residuals(fit.pca))

```
Non-constant Variance fails  
Not now
Let's try ploy transformation becuase of 
```{r}
fit.3= lm(formula = Price~Difficulty+Hash_rate+estimated_transaction_value+poly(Miners_revenue,2)+poly(Trade_volume,2)+Cost_per_transaction, data=mybitcoin)

summary(fit.3)
autoplot(fit.3, col = "goldenrod1") +  theme_minimal()
residualPlots(fit.3)

#Test for constant variance fails
lmtest::bptest(fit.3) # Breusch-Pagan test
car::ncvTest(fit.3)  # Breusch-Pagan test
shapiro.test(residuals(fit.3))
```
Residual plots looks better with Normality but Constant variance test Fails 
We will try the percent change 
```{r}
mybitcoin_final <- subset(mybitcoin, mybitcoin$Price>0)
mybitcoin_final = mutate(mybitcoin_final, pChange=(mybitcoin_final$Price-lag(mybitcoin_final$Price))/lag(mybitcoin_final$Price)*100)
mybitcoin_final = mutate(mybitcoin_final, Transaction_fees..BTC._pChange=(mybitcoin_final$Transaction_fees..BTC.-lag(mybitcoin_final$Transaction_fees..BTC.))/lag(mybitcoin_final$Transaction_fees..BTC.)*100)
mybitcoin_final = mutate(mybitcoin_final, Output_value..BTC._pChange=(mybitcoin_final$Output_value..BTC.-lag(mybitcoin_final$Output_value..BTC.))/lag(mybitcoin_final$Output_value..BTC.)*100)
mybitcoin_final = mutate(mybitcoin_final, estimated_transaction_value_pChange=(mybitcoin_final$estimated_transaction_value-lag(mybitcoin_final$estimated_transaction_value))/lag(mybitcoin_final$estimated_transaction_value)*100)
mybitcoin_final = mutate(mybitcoin_final, Miners_revenue_pChange=(mybitcoin_final$Miners_revenue-lag(mybitcoin_final$Miners_revenue))/lag(mybitcoin_final$Miners_revenue)*100)
mybitcoin_final = mutate(mybitcoin_final, No_of_transactions_pChange=(mybitcoin_final$No_of_transactions-lag(mybitcoin_final$No_of_transactions))/lag(mybitcoin_final$No_of_transactions)*100)
mybitcoin_final = mutate(mybitcoin_final, Cost_per_transaction_pChange=(mybitcoin_final$Cost_per_transaction-lag(mybitcoin_final$Cost_per_transaction))/lag(mybitcoin_final$Cost_per_transaction)*100)
mybitcoin_final = mutate(mybitcoin_final, Difficulty_pChange=(mybitcoin_final$Difficulty-lag(mybitcoin_final$Difficulty))/lag(mybitcoin_final$Difficulty)*100)
mybitcoin_final = mutate(mybitcoin_final, Hash_rate_pChange=(mybitcoin_final$Hash_rate-lag(mybitcoin_final$Hash_rate))/lag(mybitcoin_final$Hash_rate)*100)
mybitcoin_final = mutate(mybitcoin_final, Trade_volume_pChange=(mybitcoin_final$Trade_volume-lag(mybitcoin_final$Trade_volume))/lag(mybitcoin_final$Trade_volume)*100)
is.na(mybitcoin_final) = sapply(mybitcoin_final, is.infinite)
mybitcoin_final[is.na(mybitcoin_final)]=0
summary(mybitcoin_final)
```
Apply on important variable which PCA defines
```{r}
fit.pp <- lm(pChange ~ Difficulty_pChange + poly(Hash_rate_pChange,1) + poly(Miners_revenue_pChange,2) +estimated_transaction_value_pChange+Cost_per_transaction_pChange+poly(Trade_volume_pChange,2),  data=mybitcoin_final)
summary(fit.pp)

autoplot(fit.pp, col = "goldenrod1") +  theme_minimal()

#Test for constant variance fails
lmtest::bptest(fit.pp) # Breusch-Pagan test
car::ncvTest(fit.pp)  # Breusch-Pagan test
shapiro.test(residuals(fit.pp))
residualPlots(fit.pp)
```

